---
title: "PH 718 Data Management and Visualization in `R`"
subtitle: "Assignment 2"
author: Jasmine Lee (991-36-4965)
date: "`r Sys.Date()`"
output: pdf_document
bibliography: ph718.bib
link-citations: F
csl: apa-6th-edition.csl 
---

## Q1

### Answer to Q1
There are 3 cars that have 6 cylinders and manual transmission.

### Code for Q1
```{r}
data(mtcars)
df <- mtcars

# Factorize cyl
df$cyl_f <- factor(df$cyl, levels = c(4, 6, 8),
                   labels = c("4 cylinders", "6 cylinders", "8 cylinders"))

# 6 cylinders and manual transmission
cyl_6_manual <- sum(df$cyl_f == "6 cylinders" & df$am == 1)
cyl_6_manual
```

## Q2

### Answer to Q2
The average power-to-weight ratio for cars with 4 cylinders is approximately 37.925.

### Code for Q2
```{r}
data(mtcars)
df <- mtcars

# Create new variable: power_to_weight
df$power_to_weight <- df$hp / df$wt

# Compute the average power-to-weight ratio for cars with 4 cylinders
avg_ptw_4cyl <- mean(df$power_to_weight[df$cyl == 4])
avg_ptw_4cyl
```

## Q3

### Answer to Q3
There are 6 cars that are higher than 25 mpg.

### Code for Q3
```{r}
data(mtcars)
df <- mtcars

# Create new variable: high_mpg
df$high_mpg <- ifelse(df$mpg > 25, "1", "0")
# Count the number of cars with high_mpg == "1"
sum(df$high_mpg == "1")
```

## Q4

### Answer to Q4
There are 0 cars that do not satisfy the conditions.

### Code for Q4
```{r}
data(mtcars)
df <- mtcars

# Count that doesn't satisfy the conditions: 5 < mpg < 50 and hp > 0
count <- 0
for (i in 1:nrow(df)) {
  if (!((df$mpg[i] > 5 && df$mpg[i] < 50) && df$hp[i] > 0)) {
    count <- count + 1
  }
}
print(count)
```

## Q5

### Answer to Q5
The average horsepower for cars with 8 cylinders and automatic transmission is approximately 194.167.

### Code for Q5
```{r}
data(mtcars)
df <- mtcars

# Create a subset of cars with 8 cylinders and automatic transmission
cyl_8_auto <- df$cyl == 8 & df$am == 0

# Compute the mean hp for this subset
mean_hp_cyl_8_auto <- mean(df$hp[cyl_8_auto])
mean_hp_cyl_8_auto
```

## Q6

### Answer to Q6
There is only 1 car that satisfies the conditions: high_mpg == "1" and power_to_weight > 50 from the cleaned mtcars dataset.

### Code for Q6
```{r}
data(mtcars)
df <- mtcars

# Write a function: clean_mtcars
clean_mtcars <- function(df) {
  df$cyl <- factor(df$cyl, levels = c(4, 6, 8),
                   labels = c("4 cylinders", "6 cylinders", "8 cylinders"))
  df$am <- factor(df$am, levels = c(0, 1), labels = c("Automatic", "Manual"))
  df$gear <- factor(df$gear, levels = c(3, 4, 5),
                    labels = c("3 gears", "4 gears", "5 gears"))
  df$power_to_weight <- df$hp / df$wt
  df$high_mpg <- ifelse(df$mpg > 25, "1", "0")
  df
}

# Apply the function to mtcars
cleaned_df <- clean_mtcars(mtcars)

# Count cars satisfying the conditions: high_mpg == "1" and power_to_weight > 50
count <- 0
for (i in 1:nrow(cleaned_df)) {
  if (cleaned_df$high_mpg[i] == "1" && cleaned_df$power_to_weight[i] > 50) {
    count <- count + 1
  }
}
print(count)
```