---
title: "PH 718 Data Management and Visualization in `R`"
subtitle: "Assignment 3"
author: Jasmine Lee (991-36-4965)
date: "`r Sys.Date()`"
output: pdf_document
bibliography: ph718.bib
link-citations: F
csl: apa-6th-edition.csl 
---

## Q1

### Answer to Q1
There are 6580 missing values in the `selfreport` dataset.

### Code for Q1
```{r}
df <- mice::selfreport
df$hm = log(df$hm)
df$wm = log(df$wm)
sum(is.na(df))
```

## Q2

### Answer to Q2
The mean for hm and wm after imputation are approximately 5.157 and 4.333, respectively.

### Code for Q2
```{r}
df <- mice::selfreport
df$hm = log(df$hm)
df$wm = log(df$wm)

# Locations of missing values
missing_hm <- which(is.na(df$hm))
missing_wm <- which(is.na(df$wm))

# Simple imputation - To check if there's no empty data for the regression model. If we do not have missing values, no need to do imputation at all.
df$hm[missing_hm] <- mean(df$hm, na.rm = TRUE)
df$wm[missing_wm] <- mean(df$wm, na.rm = TRUE)

# Compute the mean after imputation
mean(df$hm[missing_hm])
mean(df$wm[missing_wm])
```

```{r}

```
## Q3

### Answer to Q3
For the `selfreport` dataset, there are two variables that are usually of interest for regression analysis: `hm` (height in meters) and `wm` (weight in meters).

We can exclude ID or Row Identifier, as they do not provide meaningful information for regression analysis. Additionally, any variables that are not going to be used as a predictor. Lastly, imputing an ID is non-sensical, as it is a unique identifier for each observation.

Also, since hm and wm are the variables of interest, if one of lm(wm ~ hm) or lm(hm ~ wm) is your regression model, then you can exclude the other variable as well. Those won't affect your model and they do not need to be imputed.

## Q4

### Answer to Q4
The mean of two completed datasets after imputation are adjacent to each other. Means of both completed dataset are around 5.157.

### Code for Q4
```{r}
library(mice)
df <- mice::selfreport
df$hm = log(df$hm)
df$wm = log(df$wm)

# Create 2 imputations
imp <- mice(df, m = 2, seed = 718)

# Extract the first completed dataset
data_1 <- complete(imp, 1)
# Extract the second completed dataset
data_2 <- complete(imp, 2)

# Compute the mean of hm in each
mean_hm_1 <- mean(data_1$hm)
mean_hm_2 <- mean(data_2$hm)

mean_hm_1
mean_hm_2
```