---
title: "PH 718 Data Management and Visualization in `R`"
subtitle: "Assignment 4"
author: Jasmine Lee (991-36-4965)
date: "`r Sys.Date()`"
output: pdf_document
bibliography: ph718.bib
link-citations: F
csl: apa-6th-edition.csl 
---

# Q1

## Answer to Q1(a)
The session id of 100th row of the cleaned dataset is 6397339.

## Code for Q1(a)
```{r}
library(moderndive)
library(dplyr)
library(tidyr)

ev_cleaned <- ev_charging %>%
  filter(kwh_total > 25 | distance < 10) %>%
  arrange(desc(kwh_total), distance)

ev_cleaned$session_id[100]
```

## Answer to Q1(b)
The 'cost_per_kwh' = 0 and 'hours_per_kwh' = 0.547 for the session id of 7090607.

## Code for Q1(b)
```{r}
ev_expanded <- ev_charging %>%
  mutate(
   cost_per_kwh = dollars / kwh_total,
   hours_per_kwh = charge_time_hrs / kwh_total
  )

ev_expanded %>%
  filter(session_id == 7090607) %>%
  select(cost_per_kwh, hours_per_kwh)
```

## Answer to Q1(c)
The median cost per kwh for all four facility type is 0.

And the median hours per kwh for each facility type are as follows:

- Manufacturing: 0.4225
- Office: 0.460
- Research and development: 0.500
- Other: 0.584

## Code for Q1(c)
```{r}
ev_expanded %>%
  group_by(facility_type) %>%
  summarise(
    median_cost_per_kwh = median(cost_per_kwh, na.rm = TRUE),
    median_hours_per_kwh = median(hours_per_kwh, na.rm = TRUE)
  )
```
## Answer to Q1(d)
The average kwh_total for each facility type are as follows:

- manufacturing(1): 5.753
- office(2): 5.603
- research and development(3): 5.842
- other(4): 7.217

## Code for Q1(d)
```{r}
ev_expanded_2 <- ev_expanded %>%
  mutate(
    facility_number = case_when(
      facility_type == "manufacturing" ~ 1,
      facility_type == "office" ~ 2,
      facility_type == "research and development" ~ 3,
      facility_type == "other" ~ 4
    )
  )

ev_expanded_2 %>%
  group_by(facility_number) %>%
  summarise(avg_kwh_total = mean(kwh_total, na.rm = TRUE))
```

# Q2

## Answer to Q2(a)
The average for waiting and duration grouped by existence of webcam are approximately 218.500 and 217.380, respectively.

## Code for Q2(a)
```{r}
library(moderndive)
library(dplyr)
library(tidyr)

faithful_summary <- old_faithful_2024 %>%
  group_by(webcam) %>%
  summarise(
    avg_waiting = mean(waiting, na.rm = TRUE),
    avg_duration = mean(duration, na.rm = TRUE)
  )

faithful_summary
```
## Answer to Q2(b)
The average for waiting and duration grouped by existence of webcam are approximately as follows:

- Webcam: No, Variable: avg_waiting = 172.333
- Webcam: No, Variable: avg_duration = 218.500
- Webcam: Yes, Variable: avg_waiting = 159.472
- Webcam: Yes, Variable: avg_duration = 217.380

## Code for Q2(b)
```{r}
faithful_summary_long <- faithful_summary %>%
  pivot_longer(
    cols = c(avg_waiting, avg_duration),
    names_to = "Variable",
    values_to = "Value"
  )

faithful_summary_long
```
## Answer to Q2(c)
Nothing changes since we only changed the entry of the Variable column.

## Code for Q2(c)
```{r}
faithful_summary_long <- faithful_summary_long %>%
  mutate(
    Variable = sub("avg_", "", Variable)
  )

faithful_summary_long
```