---
title: "PH718_Part6"
author: "Jasmine"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r, echo = T, include = T, eval = T, cache = T, message = T}
college = read.csv(file = "https://www.statlearning.com/s/College.csv", header = T)

rownames(college) <- college[, 1]
college <- college[, -1]
```

```{r, echo = T, include = T, eval = F, cache = T, message = T}
summary(college)
```

## Exploring relationships via `plot()`
```{r, echo = T, include = T, eval = T, cache = T, message = T}
plot(
  college$Apps, college$Accept,
  xlab = "Number of Applications",
  ylab = "Number of Acceptances",
  main = "Applications vs Acceptances",
  col = "cyan",
  pch = 19
)
abline(lm(college$Accept ~ college$Apps), col = "green")
```

## Visualizing many relationships: pairs()
```{r, echo = T, include = T, eval = T, cache = T, message = T}
pairs(college[, 2:10])
```

## Modern alternative to `pairs()`: `GGally::ggpairs()`
```{r, echo = T, include = T, eval = T, cache = T, message = T}
# college <- college[, -1] # this might delete the 'Private' column, the first column, so be careful
GGally::ggpairs(college[, 2:10])
```

## Comparing groups via `boxplot()`
```{r, echo = T, include = T, eval = T, cache = T, message = T}
college$Private <- factor(college$Private)
boxplot(
  Outstate ~ Private,
  data = college,
  col = c("cyan", "green"),
  main = "Out-of-State Tuition by College Type",
  ylab = "Out-of-State Tuition"
)
```

```{r, echo = T, include = T, eval = T, cache = T, message = T}
Elite <- rep("No", nrow(college))
Elite[college$Top10perc > 50] <- "Yes"
Elite <- as.factor(Elite)
college <- data.frame(college, Elite) # or college$Elite = Elite
```

```{r, echo = T, include = T, eval = T, cache = T, message = T}
boxplot(Outstate ~ Elite, data = college)
```

```{r, echo = T, include = T, eval = F, cache = T, message = T}
hist(
college$Top25perc,
main = "Distribution of Top 25% Students",
xlab = "Percent of Students in Top 25%"
)
```

```{r, echo = T, include = T, eval = T, cache = T, message = T}
hist(
college$Top25perc,
main = "Distribution of Top 25% Students",
xlab = "Percent of Students in Top 25%"
)
```

```{r, echo = T, include = T, eval = T, cache = T, message = T}
par(mfrow = c(1, 2))
hist(college$Top25perc, breaks = 1, freq = T, main = 'breaks=1')
hist(college$Top25perc, breaks = 1, freq = F, main = 'breaks=1')
hist(college$Top25perc, breaks = 10, freq = T, main = 'breaks=10')
hist(college$Top25perc, breaks = 100, freq = T, main = 'breaks=100')
hist(college$Top25perc, breaks = 1000, freq = T, main = 'breaks=1000')
par(mfrow = c(1, 1))
```

## Heatmaps of correlation matirx via image()
```{r, echo = T, include = T, eval = T, cache = T, message = T}
num_var_idx = NULL
for (j in 1:ncol(college)) {
  if(is.numeric(college[, j])) {
    num_var_idx <- c(num_var_idx, j)
  }
}
college_num <- college[, num_var_idx] # subsetting numeric variables
cor_matrix <- cor(college_num)
col_palette <- colorRampPalette(c("maroon","seagreen2","yellow"))(100)
image(
  1:ncol(cor_matrix),
  1:ncol(cor_matrix),
  cor_matrix,
  axes = FALSE,
  xlab = "",
  ylab = "",
  main = "Heatmap of Covariance Matrix",
  col = col_palette
)
axis(1, at = 1:ncol(cor_matrix), labels = colnames(cor_matrix), las = 2)
axis(2, at = 1:ncol(cor_matrix), labels = colnames(cor_matrix), las = 2)
```

## Interpretation:
<!-- # Need this line to be spaced to print the list -->
- Colors close to green or red indicate larger correlation coefficients.
- Large positive values suggest variables increase together.
- Neg. values indicate variables move in opposite directions.